\textbf{Using generalized additive models to analyze biomedical non-linear longitudinal data}

### 2021-06-09 Manuscripts/Manuscript by chapters -v2/04-Longitudinal analysis with GAMs.Rmd
Made changes suggested by JRT

The simulated data is based on the reported longitudinal changes in oxygen saturation ($\mbox{StO}_2$) in subcutaneous tumors that appear in Figure 3, C in [@vishwanath2009]. In the paper, diffuse reflectance spectroscopy was used to quantify $\mbox{StO}_2$ changes in both groups at the same time points (days 0, 2, 5, 7 and 10). In the "Treatment" group (chemotherapy) an increase in $\mbox{StO}_2$ is observed through time, while a decrease is seen in the "Control" (saline) group. Following the reported trend, we simulated 10 normally distributed observations at each time point with a standard deviation (SD) of 10% (matching the SD in the original paper). 
The simulated and real data appear in Figure \@ref(fig:sim-smooth-plot), A and the inlet, respectively.

@@

`m1 <- gam(StO2_sim ~ Group + s(Day, by=Group, k=5), method='REML',data=dat_sim)`

This syntax specifies that `m1` will store the model, and that the change in the simulated oxygen saturation (`StO2_sim`) is modeled using independent smooths for `Group` and `Day` (the parenthesis preceded by `s`) using 5 knots. The smooth is constructed by default using thin plate regression splines, but other splines can be used if desired, including Gaussian process smooths [@simpson2018]. The parametric term `Group` is added to quantify differences in the effect of treatment between groups, and the `method` chosen to estimate the smoothing parameters is the restricted maximum likelihood (REML) [@wood2017]. When the smooths are plotted over the raw data, it is clear that the model has been able to capture the trend of the change of $\mbox{StO}_2$ for each group across time (Figure \@ref(fig:sim-smooth-plot),B). Model diagnostics can be obtained using the `gam.check` function, and the function `appraise` from the package _gratia_ [@gratia]. A guide for model selection and diagnostics is in the Appendix, and an in-depth analysis can be found in [@wood2017] and [@harezlak2018].

One question that might arise at this point is "what is the fit that an rm-ANOVA model produces for the simulated data?". The rm-ANOVA model, which corresponds to Equation \@ref(eq:linear-model) is presented in Figure \@ref(fig:sim-smooth-plot), C. This is a typical case of model misspecification: The slopes of each group are different, which would lead to a _p-value_ indicating significance for the treatment and time effects, but the model is not capturing the changes that occur at days 2 and between days 5 and 7, whereas the GAM model is able to do so (Figure \@ref(fig:sim-smooth-plot), B) .

@@

Because GAMs do not require equally-spaced or complete observations for all subjects, they are advantageous to analyze longitudinal data where missingness exists. The rationale behind this is that GAMs are able to pick the trend in the data even when some observations are missing. However, this usually causes the resulting smooths to have wider confidence intervals and less ability to pick certain trends. Consider the  simulated $\mbox{StO}_2$ values from Figure (\@ref(fig:sim-smooth-plot), B). If 40% of the total observations are randomly deleted and the same interaction GAM fitted for the complete dataset is used, the resulting smooths are still able to show a different trend for each group, but it can be seen that the smooths overlap during the first 3 days because with less data points, the trend is less pronounced than in the full dataset (\@ref(fig:sim-smooth-plot), D). Although the confidence intervals have increased for both smooths, the model still shows different trends with as little as 4 observations per group at certain time points.

@@

(ref:sim-plot-caption) Simulated data and smooths for oxygen saturation in tumors. A: Simulated data that follows previously reported trends (inset) in tumors under chemotherapy (Treatment) or saline (Control) treatment. Simulated data is from  a normal distribution with standard deviation of 10% with 10 observations per time point. Lines indicate mean oxygen saturation B: Smooths from the GAM model for the full simulated data with interaction of Group and Treatment. Lines represent trends for each group, shaded regions are 95% confidence intervals. C: rm-ANOVA model for the simulated data, the model does not capture the changes in each group over time. D: Smooths for the GAM model for the simulated data with 40% of its observations missing. Lines represent trends for each group, shaded regions are 95% confidence intervals.

@@

my_list<-pairwise_limits(comp_StO2_full)

c1<-ggplot(comp_StO2_full, aes(x = Day, y = diff, group = pair)) +
    annotate("rect",
                xmin =my_list$init1, xmax =my_list$final1,ymin=-Inf,ymax=Inf,
                fill='#30123BFF',
                alpha = 0.5,
                ) +
  annotate("text",
             x=1.5,
             y=-18,
             label="Control>Treatment",
           size=8,
           angle=90
           )+
    annotate("rect",
             xmin =my_list$init2, xmax =my_list$final2,ymin=-Inf,ymax=Inf,
             fill='#7A0403FF',
             alpha = 0.5,
    ) +
  annotate("text",
             x=6,
             y=-18,
             label="Treatment>Control",
             size=8,
           angle=90
           )+
    geom_ribbon(aes(ymin = lower, ymax = upper),
                alpha = 0.5,
                fill='#DB3A07FF') +
    geom_line(data=comp_StO2_full,aes(y=0),size=0.5)+
    geom_line(color='black',size=1) +

@@

c2<-ggplot(comp_StO2_missing, aes(x = Day, y = diff, group = pair)) +
    annotate("rect",
             xmin =my_list$init1, xmax =my_list$final1,ymin=-Inf,ymax=Inf,
             fill='#30123BFF',
             alpha = 0.5
             ) +
  annotate("text",
             x=1.5,
             y=-18,
             label="Control>Treatment",
           size=8,
           angle=90
           )+
    annotate("rect",
             xmin =my_list$init2, xmax =my_list$final2,ymin=-Inf,ymax=Inf,
             fill='#7A0403FF',
             alpha = 0.5,
    ) +
  annotate("text",
             x=6,
             y=-18,
             label="Treatment>Control",
             size=8,
           angle=90
           )+
    geom_ribbon(aes(ymin = lower, ymax = upper),
                alpha = 0.5,
                fill='#DB3A07FF') +
    geom_line(data=comp_StO2_missing,aes(y=0),size=0.5)+
    geom_line(color='black',size=1) +
    facet_wrap(~ pair) +
    theme_classic()+
    labs(x = 'Days', y = expression(paste('Difference in StO'[2] )))+
    scale_x_continuous(breaks=c(0,2,5,7,10))+
    theme(
        text=element_text(size=18),
        legend.title=element_blank()
    )

@@

(ref:pairwise-comp) Pairwise comparisons for smooth terms. A: Pairwise comparisons for the full dataset.  B: Pairwise comparisons for the dataset with missing observations. Significant differences exist where the interval does not cover 0. In both cases the effect of treatment is significant after day 3.

@@

At the core of a biomedical longitudinal study lies the question of a significant difference between the effect of two or more treatments in different groups. Whereas in rm-ANOVA a _post-hoc_ analysis is required to answer such question by calculating some _p-values_ after multiple comparisons, GAMs can use a different approach to estimate significance. In essence, the idea behind the estimation of significance in GAMs across different treatment groups is that if the _difference_ between the empirical Bayesian confidence intervals of the fitted smooths for such groups is  non-zero, then a significant difference exists at that time point(s). The absence of a _p-value_ in this case might seem odd, but the empirical Bayesian confidence interval interpretation can be conceptualized in the following manner: Different trends in each group are an indication of an effect by the treatment. This is what happens for the simulated data in Figure \@ref(fig:sim-smooth-plot), A where the chemotherapy causes $\mbox{StO}_2$ to increase over time.

@@

Consider the calculation of pairwise differences for the smooths in Figure \@ref(fig:sim-smooth-plot), B and D. Figure \@ref(fig:plot-pairwise-comp), shows the comparison between each treatment group for the full and missing datasets. Here, the  "Control" group is used as the reference to which "Treatment" group is being compared. Of notice, the pairwise comparison has been set on the response scale (see Appendix for code details), because otherwise the comparison appears shifted and is not intuitively easy to relate to the original data.

With this correction in mind, the shaded regions over the confidence interval (where it does not cover 0) indicate the time interval where each group has a higher effect than the other. Notice that the shaded region between days 0 and $\approx$ 2 for the full dataset indicates that through that time, the "Control" group has higher $\mbox{StO}_2$, but as therapy progresses the effect is reversed and by $\approx$ 3 day  it is the "Treatment" group the one that has greater $\mbox{StO}_2$. This would suggest that the effect of chemotherapy in the "Treatment" group becomes significant after day 3 for the model used. Moreover, notice that although there is no actual measurement at day 3, the model is capable of providing an estimate of when the shift in $\mbox{StO}_2$ occurs.

On the data with missing observations (Figure \@ref(fig:sim-smooth-plot), D), the confidence intervals of the smooths overlap between days 0 and 3. Consequently, the smooth pairwise comparison (Figure \@ref(fig:plot-pairwise-comp), B) shows that there is not a significant difference between the groups during that period, but is still able to pick the change on day 3 as the full dataset smooth pairwise comparison.

>>>>

The simulated data is based on the reported longitudinal changes in oxygen saturation ($\mbox{StO}_2$) in subcutaneous tumors that appear in Figure 3C in [@vishwanath2009]. In the paper, diffuse reflectance spectroscopy was used to quantify $\mbox{StO}_2$ changes in both groups at the same time points (days 0, 2, 5, 7 and 10). In the "Treatment" group (chemotherapy) an increase in $\mbox{StO}_2$ is observed through time, while a decrease is seen in the "Control" (saline) group. Following the reported trend, we simulated 10 normally distributed observations at each time point with a standard deviation (SD) of 10% (matching the SD in the original paper). 
The simulated and real data appear in Figure \@ref(fig:sim-smooth-plot)A and the inset, respectively.

@@

`m1 <- gam(StO2_sim ~ Group + s(Day, by=Group, k=5), method='REML, data = dat_sim)`

This syntax specifies that `m1` will store the model, and that the change in the simulated oxygen saturation (`StO2_sim`) is modeled using independent smooths over `Day` for each `Group` (the parenthesis preceded by `s`) using 5 knots. The smooth is constructed by default using thin plate regression splines, but other splines can be used if desired, including Gaussian process smooths [@simpson2018]. The parametric term `Group` is added to quantify overall mean differences in the effect of treatment between groups, and the `method` chosen to estimate the smoothing parameters is the restricted maximum likelihood (REML) [@wood2017]. When the smooths are plotted over the raw data, it is clear that the model has been able to capture the trend of the change of $\mbox{StO}_2$ for each group across time (Figure \@ref(fig:sim-smooth-plot)B). Model diagnostics can be obtained using the `gam.check` function, and the function `appraise` from the package _gratia_ [@gratia]. A guide for model selection and diagnostics is in the [Appendix](#workflow), and an in-depth analysis can be found in [@wood2017] and [@harezlak2018].

One question that might arise at this point is "what is the fit that an rm-ANOVA model produces for the simulated data?". The rm-ANOVA model, which corresponds to Equation \@ref(eq:linear-model) is presented in Figure \@ref(fig:sim-smooth-plot)C. This is a typical case of model misspecification: The slopes of each group are different, which would lead to a _p-value_ indicating significance for the treatment and time effects, but the model is not capturing the changes that occur at days 2 and between days 5 and 7, whereas the GAM model is able to reliably estimate the trend over all timepoints (Figure \@ref(fig:sim-smooth-plot)B) .

@@

Because GAMs do not require equally-spaced or complete observations for all subjects, they are advantageous to analyze longitudinal data where missingness exists. The rationale behind this is that GAMs are able to pick the trend in the data even when some observations are missing. However, this usually causes the resulting smooths to have wider confidence intervals and less ability to pick certain trends. Consider the  simulated $\mbox{StO}_2$ values from Figure \@ref(fig:sim-smooth-plot)B. If 40% of the total observations are randomly deleted and the same interaction GAM fitted for the complete dataset is used, the resulting smooths are still able to show a different trend for each group, but because the empirical Bayesian credible intervals for the smooths overlap during the first 3 days with fewer data points, the trend is less pronounced than in the full dataset (Figure \@ref(fig:sim-smooth-plot)D). Although the confidence intervals have increased for both smooths, the model still shows different trends with as little as 4 observations per group at certain time points.

@@

(ref:sim-plot-caption) Simulated data and smooths for oxygen saturation in tumors. A: Simulated data that follows previously reported trends (inset) in tumors under chemotherapy (Treatment) or saline (Control) treatment. Simulated data is from  a normal distribution with standard deviation of 10% with 10 observations per time point. Lines indicate mean oxygen saturation B: Smooths from the GAM model for the full simulated data with interaction of Group and Treatment. Lines represent trends for each group, shaded regions are 95% confidence intervals. C: The rm-ANOVA model for the simulated data, which does not capture the changes in each group over time. D: Smooths for the GAM model for the simulated data with 40% of its observations missing. Lines represent trends for each group, shaded regions are 95% empirical Bayesian confidence intervals.

@@

my_list<-pairwise_limits(comp_StO2_full)
rib_col<-'#EDD03AFF' #color for ribbon

c1<-ggplot(comp_StO2_full, aes(x = Day, y = diff, group = pair)) +
    annotate("rect",
                xmin =my_list$init1, xmax =my_list$final1,ymin=-Inf,ymax=Inf,
                fill='#30123BFF',
                alpha = 0.5,
                ) +
  annotate("text",
             x=1.5,
             y=-18,
             label="Control>Treatment",
           size=8,
           angle=90
           )+
    annotate("rect",
             xmin =my_list$init2, xmax =my_list$final2,ymin=-Inf,ymax=Inf,
             fill='#7A0403FF',
             alpha = 0.5,
    ) +
  annotate("text",
             x=6,
             y=-18,
             label="Treatment>Control",
             size=8,
           angle=90
           )+
    geom_ribbon(aes(ymin = lower, ymax = upper),
                alpha = 0.5,
                fill=rib_col) +
    geom_line(data=comp_StO2_full,aes(y=0),size=0.5)+
    geom_line(color='black',size=1) +

@@

c2<-ggplot(comp_StO2_missing, aes(x = Day, y = diff, group = pair)) +
    annotate("rect",
             xmin =my_list$init1, xmax =my_list$final1,ymin=-Inf,ymax=Inf,
             fill='#30123BFF',
             alpha = 0.5
             ) +
  annotate("text",
             x=1.5,
             y=-18,
             label="Control>Treatment",
           size=8,
           angle=90
           )+
    annotate("rect",
             xmin =my_list$init2, xmax =my_list$final2,ymin=-Inf,ymax=Inf,
             fill='#7A0403FF',
             alpha = 0.5,
    ) +
  annotate("text",
             x=6,
             y=-18,
             label="Treatment>Control",
             size=8,
           angle=90
           )+
    geom_ribbon(aes(ymin = lower, ymax = upper),
                alpha = 0.5,
                fill=rib_col) +
    geom_line(data=comp_StO2_missing,aes(y=0),size=0.5)+
    geom_line(color='black',size=1) +
    facet_wrap(~ pair) +
    theme_classic()+
    labs(x = 'Days', y = expression(paste('Difference in StO'[2] )))+
    scale_x_continuous(breaks=c(0,2,5,7,10))+
    theme(
        text=element_text(size=18),
        legend.title=element_blank()
    )

@@

(ref:pairwise-comp) Pairwise comparisons for smooth terms. A: Pairwise comparisons for the full dataset.  B: Pairwise comparisons for the dataset with missing observations. Significant differences exist where the 95% empirical Bayesian credible interval does not cover 0. In both cases the effect of treatment is significant after day 3.

@@

At the core of a biomedical longitudinal study lies the question of a significant difference between the effect of two or more treatments in different groups. Whereas in rm-ANOVA a _post-hoc_ analysis is required to answer such question by calculating some _p-values_ after multiple comparisons, GAMs can use a different approach to estimate significance. In essence, the idea behind the estimation of significance in GAMs across different treatment groups is that if the _difference_ between the empirical Bayesian confidence intervals of the fitted smooths for such groups is  non-zero, then a significant difference exists at that time point(s). The absence of a _p-value_ in this case might seem odd, but the empirical Bayesian confidence interval interpretation can be conceptualized in the following manner: Different trends in each group are an indication of an effect by the treatment. This is what happens for the simulated data in Figure \@ref(fig:sim-smooth-plot)A, where the chemotherapy causes $\mbox{StO}_2$ to increase over time.

@@

Consider the calculation of pairwise differences for the smooths in Figure \@ref(fig:sim-smooth-plot)B and Figure \@ref(fig:sim-smooth-plot)D. Figure \@ref(fig:plot-pairwise-comp) shows the comparison between each treatment group for the full and missing datasets. Here, the  "Control" group is used as the reference to which "Treatment" group is being compared. Of notice, the pairwise comparison has been set on the response scale (see Appendix for code details), because otherwise the comparison appears shifted and is not intuitively easy to relate to the original data.

With this correction in mind, the shaded regions over the confidence interval (where it does not cover 0) indicate the time interval where each group has a higher effect than the other. Notice that the shaded region between days 0 and $\approx$ 2 for the full dataset indicates that through that time, the "Control" group has higher mean $\mbox{StO}_2$, but as therapy progresses the effect is reversed and by $\approx$ 3 day  it is the "Treatment" group the one that on average, has greater $\mbox{StO}_2$. This would suggest that the effect of chemotherapy in the "Treatment" group becomes significant after day 3 for the given model. Moreover, notice that although there is no actual measurement at day 3, the model is capable of providing an estimate of when the shift in mean $\mbox{StO}_2$ occurs.

On the data with missing observations (Figure \@ref(fig:sim-smooth-plot)D), the empirical Bayesian credible confidence intervals of the smooths overlap between days 0 and 3. Consequently, the smooth pairwise comparison (Figure \@ref(fig:plot-pairwise-comp)B) shows that there is no evidence of a significant difference between the groups during that period, but is still able to pick the change on day 3 as the full dataset smooth pairwise comparison.

### 2021-06-09 Manuscripts/Manuscript by chapters -v2/07-Appendix.Rmd
Made changes suggested by JRT

## A basic Workflow for GAMs

@@

    my_list<-pairwise_limits(comp_StO2)

#plot the difference
c1<-ggplot(comp_StO2, aes(x = Day, y = diff, group = pair)) +
  #shaded region
  annotate("rect",
                xmin =my_list$init1, xmax =my_list$final1,ymin=-Inf,ymax=Inf,
                fill='#30123BFF',
                alpha = 0.5,
                ) +
  annotate("text",
             x=1.5,
             y=-10,
             label="Control",size=10
           )+
  #shaded region  
  annotate("rect",
             xmin =my_list$init2, xmax =my_list$final2,ymin=-Inf,ymax=Inf,
             fill='#7A0403FF',
             alpha = 0.5
           ) +
  annotate("text",
             x=6,
             y=-10,
             label="Treatment",
             size=10
           )+
  #ribbon for difference confidence interval  
  geom_ribbon(aes(ymin = lower, ymax = upper),
                alpha = 0.5,
                fill='#DB3A07FF') +
    geom_line(color='black',size=1) +
    geom_line(data=comp_StO2,aes(y=0),size=0.5)+
    facet_wrap(~ pair) +
    theme_classic()+
    labs(x = 'Days', y = expression(paste('Difference in StO'[2] )))+
    scale_x_continuous(breaks=c(0,2,5,7,10))+
    theme(
        text=element_text(size=18),
        legend.title=element_blank()
    )

@@

This section covers the code used to generate Figure \@ref(fig:sim-smooth-plot), where the simulated data, fit of the "final" GAM (`m1`), linear model  and GAM on data with missing observations are presented. Note that panel A in Figure \@ref(fig:sim-smooth-plot) and the inlet are generated in the code chunk where the data is simulated in Section \@ref(tumor-data-simulation), and are called later to build the figure.

@@

my_list<-pairwise_limits(comp_StO2_full)

c1<-ggplot(comp_StO2_full, aes(x = Day, y = diff, group = pair)) +
    annotate("rect",
                xmin =my_list$init1, xmax =my_list$final1,ymin=-Inf,ymax=Inf,
                fill='#30123BFF',
                alpha = 0.5,
                ) +
  annotate("text",
             x=1.5,
             y=-15,
             label="Control",
           size=10,
           angle=90
           )+
    annotate("rect",
             xmin =my_list$init2, xmax =my_list$final2,ymin=-Inf,ymax=Inf,
             fill='#7A0403FF',
             alpha = 0.5,
    ) +
  annotate("text",
             x=6,
             y=-15,
             label="Treatment",
             size=10,
           angle=90
           )+
    geom_ribbon(aes(ymin = lower, ymax = upper),
                alpha = 0.5,
                fill='#DB3A07FF') +
    geom_line(data=comp_StO2_full,aes(y=0),size=0.5)+
    geom_line(color='black',size=1) +

@@

c2<-ggplot(comp_StO2_missing, aes(x = Day, y = diff, group = pair)) +
    annotate("rect",
             xmin =my_list$init1, xmax =my_list$final1,ymin=-Inf,ymax=Inf,
             fill='#30123BFF',
             alpha = 0.5,
    ) +
  annotate("text",
             x=1.5,
             y=-15,
             label="Control",size=10
           )+
    annotate("rect",
             xmin =my_list$init2, xmax =my_list$final2,ymin=-Inf,ymax=Inf,
             fill='#7A0403FF',
             alpha = 0.5,
    ) +
  annotate("text",
             x=6,
             y=-15,
             label="Treatment",
             size=10)+
    geom_ribbon(aes(ymin = lower, ymax = upper),
                alpha = 0.5,
                fill='#DB3A07FF') +
    geom_line(data=comp_StO2_missing,aes(y=0),size=0.5)+
    geom_line(color='black',size=1) +
    facet_wrap(~ pair) +
    theme_classic()+
    labs(x = 'Days', y = expression(paste('Difference in StO'[2] )))+
    scale_x_continuous(breaks=c(0,2,5,7,10))+
    theme(
        text=element_text(size=18),
        legend.title=element_blank()
    )

>>>>

## A basic Workflow for GAMs{#workflow}

@@

    my_list<-pairwise_limits(comp_StO2)
  rib_col<-'#EDD03AFF' #color for the ribbon
#plot the difference
c1<-ggplot(comp_StO2, aes(x = Day, y = diff, group = pair)) +
  #shaded region
  annotate("rect",
                xmin =my_list$init1, xmax =my_list$final1,ymin=-Inf,ymax=Inf,
                fill='#30123BFF',
                alpha = 0.5,
                ) +
  annotate("text",
             x=1.5,
             y=-10,
             label="Control",size=10
           )+
  #shaded region  
  annotate("rect",
             xmin =my_list$init2, xmax =my_list$final2,ymin=-Inf,ymax=Inf,
             fill='#7A0403FF',
             alpha = 0.5
           ) +
  annotate("text",
             x=6,
             y=-10,
             label="Treatment",
             size=10
           )+
  #ribbon for difference confidence interval  
  geom_ribbon(aes(ymin = lower, ymax = upper),
                alpha = 0.5,
                fill=rib_col) +
    geom_line(color='black',size=1) +
    geom_line(data=comp_StO2,aes(y=0),size=0.5)+
    facet_wrap(~ pair) +
    theme_classic()+
    labs(x = 'Days', y = expression(paste('Difference in StO'[2] )))+
    scale_x_continuous(breaks=c(0,2,5,7,10))+
    theme(
        text=element_text(size=18),
        legend.title=element_blank()
    )

@@

This section covers the code used to generate Figure \@ref(fig:sim-smooth-plot), where the simulated data, fit of the "final" GAM (`m1`), linear model  and GAM on data with missing observations are presented. Note that panel A in Figure \@ref(fig:sim-smooth-plot) and the inset are generated in the code chunk where the data is simulated in Section \@ref(tumor-data-simulation), and are called later to build the figure.

@@

my_list<-pairwise_limits(comp_StO2_full)
rib_col<-'#EDD03AFF'

c1<-ggplot(comp_StO2_full, aes(x = Day, y = diff, group = pair)) +
    annotate("rect",
                xmin =my_list$init1, xmax =my_list$final1,ymin=-Inf,ymax=Inf,
                fill='#30123BFF',
                alpha = 0.5,
                ) +
  annotate("text",
             x=1.5,
             y=-18,
             label="Control>Treatment",
           size=8,
           angle=90
           )+
    annotate("rect",
             xmin =my_list$init2, xmax =my_list$final2,ymin=-Inf,ymax=Inf,
             fill='#7A0403FF',
             alpha = 0.5,
    ) +
  annotate("text",
             x=6,
             y=-18,
             label="Treatment>Control",
             size=8,
           angle=90
           )+
    geom_ribbon(aes(ymin = lower, ymax = upper),
                alpha = 0.5,
                fill=rib_col) +
    geom_line(data=comp_StO2_full,aes(y=0),size=0.5)+
    geom_line(color='black',size=1) +

@@

c2<-ggplot(comp_StO2_missing, aes(x = Day, y = diff, group = pair)) +
    annotate("rect",
             xmin =my_list$init1, xmax =my_list$final1,ymin=-Inf,ymax=Inf,
             fill='#30123BFF',
             alpha = 0.5,
    ) +
  annotate("text",
             x=1.5,
             y=-18,
             label="Control>Treatment",
           size=8
           )+
    annotate("rect",
             xmin =my_list$init2, xmax =my_list$final2,ymin=-Inf,ymax=Inf,
             fill='#7A0403FF',
             alpha = 0.5,
    ) +
  annotate("text",
             x=6,
             y=-18,
             label="Treatment>Control",
             size=8)+
    geom_ribbon(aes(ymin = lower, ymax = upper),
                alpha = 0.5,
                fill=rib_col) +
    geom_line(data=comp_StO2_missing,aes(y=0),size=0.5)+
    geom_line(color='black',size=1) +
    facet_wrap(~ pair) +
    theme_classic()+
    labs(x = 'Days', y = expression(paste('Difference in StO'[2] )))+
    scale_x_continuous(breaks=c(0,2,5,7,10))+
    theme(
        text=element_text(size=18),
        legend.title=element_blank()
    )
